---
title: "class_11_21_lubridate"
author: "lisa liubovich"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("lubridate")
library(lubridate)
```

> # **Lubridate**
>
> stringr deals with strings
>
> forcat deals with factors
>
> lubridate is to work with date
>
> what kind of classes/data objects when you're working with time? 4 diff kinds
>
> -   date: just for date
>
> -   posixct: date and time and timezone
>
> -   

```{r}
today() # this is date
now() # this is posixct
```

> there is no built in R function to find the current time without the date
>
> use hms::ashms(now() to get the current time

```{r}
hms::as_hms(now())
class(hms::as_hms(now())) # be careful not to mix difftime with date_time, etc --> stick to one format if possible
```

> use difftime() and as.difftime() to create time difference object

```{r}
difftime(Sys.time(), make_datetime(month = 9, year = 2001, day = 11) # this is a posixct, sys.time are posixct
         units = "days") # creates a time object that calculates the time difference between 9/11 and now (lol why)
```

```{r}
as.difftime(c("0:3:20", "11:23:15")) # works differently, converts vectors of strings to time and compute them based on means

# if you want to find the difference of times, you need to assign this to a variable
```

```{r}
t <- as.difftime(c("0:3:20", "11:23:15"))
difftime1 <- t[2] - t[1]
difftime1

# this computes the difference by assinging; remember that this is basically just parsing
```

> if you want to just have the time alone, use make_datetime() and only work with the appropriate arguments

## parsing dates and times

> we used readr for this in the past but there was one main issue: you will run into errors if things aren't parsed in the way r expects

```{r}
x <- parse_date("10/11/2023", format = "%m/%d/%Y") # must correctly match format to avoid parsing problems
x
```

```{r}
class(x)
```

## parsing using lubridate

> lubridate makes parsing easier relative to readr
>
> naming: ymd --\> y stands for year, m stands for month, d stands for day

```{r}
mdy(c("01/10/2023", "January, 1 2023")) # order of the input must match the function
```

> ms(), hm(), and hms() do not recognize dash (-) as a separator because they treat it as negative time. so here you must use parse_time.

```{r}
# hms("043450") # can't detect which number is hour, minute, second
hms("04 34 50") # has to be separated somehow
```

```{r}
class(ymd_hms("20231123 04 37/20"))
```

### **exercise**

```{r}
d4 <- c("August 19 (2015)", "July 1 (2015)")
mdy(d4) # mdy matches the order of the date; much simpler than readr
```

## creating date-time values with individual components

```{r}
install.packages("nycflights13")
library(nycflights13)
```

```{r}
# head(flights)
# lets recreate this using lubridate
# flights %>% 
  # mutate(datetime = make_datetime(year = year,
                                 # month = month,
                                 # day = day,
                                  #hour = hour,
                                 # minute = minute)) ->
# flights
# select(flights, datetime)


flights %>% 
  mutate(datetime = ymd_hm(paste(year, month, day, hour, minute))) -> flights
select(flights, datetime)
```

> having it in this format makes them much easier to work with
>
> it will also make it easier to filter by date using as_date() and ymd --\> **this is important and will show up on the quiz**

```{r THIS IDEA WILL BE IMPORTANT!!!!}
flights %>% 
  filter(as_date(datetime) == ymd("20130704")) %>% # outputs only the date of this datetime; now we can ask "are these dates equal to each other?" and filter works properly
  ggplot(aes(x = datetime)) +
  geom_freqpoly(binwidth = 600)
```

# **Use `as_date()` and `as_datetime()` to Coerce Objects to Dates/Times**

> -   I used `as_date()` in the previous example. This function will *try* to coerce an object to a date. Sometimes successfully! It is particularly useful for extracting the date component of a `POSIXct` object.
>
> -   `as_datetime()` tries to coerce an object to a `POSIXct` object.

### exercise

> Create a date variable from the following data frame. Then filter out all rows before Feb 1, 2010. If you finish early, try to repeat by converting the month variable to a numeric representation of the month without using month_num.

```{r}
fake <- tribble(~year, ~month, ~day, ~month_num,
                ##----/-------/----------------
                2018,  "Oct",  1,    10,
                2011,  "Nov",  2,    11,
                2019,  "Dec",  3,    12,
                2010,  "JAN",  5,     1,
                1999,  "MAr",  1,     3,
                1987,  "ApR",  3,     4,
                2020,  "maY",  2,     5,
                2010,  "May",  4,     5)
```

```{r}
fake %>% 
  mutate(date = ymd(paste(year, month_num, day))) %>%  # paste puts the characters in the string together, kind of like copy paste
  filter(date > "2011/02/01")
```

```{r}
fake %>% 
  mutate(date = make_date(year = year, month = month_num, day = day)) %>% 
  filter(date > "2011/02/01")
```

# **{lubridate} has Functions for Getting/Setting Components of a date-time**

## **Getting or Extracting the Component of a date-time**

> important functions to memorize for extracting information:
>
> -   `year()` extracts the year.
>
> -   `month()` extracts the month.
>
> -   `week()` extracts the week.
>
> -   `mday()` extracts the day of the month (1, 2, 3, ...).
>
> -   `wday()` extracts the day of the week (Saturday, Sunday, Monday ...).
>
> -   `yday()` extracts the day of the year (1, 2, 3, ...)
>
> -   `hour()` extracts the hour.
>
> -   `minute()` extract the minute.
>
> -   `second()` extracts the second.

```{r}
ddat <- mdy_hms("01/02/1970 03:51:44")
ddat
```

```{r}
year(ddat)
```

```{r}
month(ddat, label = TRUE) # be careful about classes so that you don't run into errors; check if your classes are consistent with what you want
```

```{r}
week(ddat)
```

```{r}
mday(ddat)
```

```{r}
wday(ddat, label = TRUE)
```

```{r}
yday(ddat)
```

## **Setting the Component of a date-time**

> You can use the same functions to set or overwrite components of a date-time.

```{r}
ddat <- mdy_hms("01/02/1970 03:51:44")
ddat
```

```{r}
year(ddat) <- 1988
ddat
```

## **Rounding Date-Times**

> You can round to the levels of individual components

```{r}
ddat <- mdy_hms("01/04/1970 03:51:44")
ddat
```

```{r}
round_date(ddat, unit = "year")
```

```{r}
 round_date(ddat, unit = "month")
```

```{r}
 round_date(ddat, unit = "day")
```

```{r}
 round_date(ddat, unit = "hour")
```

> For rounding date-times exactly halfway between two consecutive units, the convention is to round up.

```{r}
ddat <- mdy_hms("01/04/1970 12:30:00")
 round_date(ddat, unit = "hour")
```

```{r}
 ddat <- mdy_hms("02/14/1970 12:00:00")
 round_date(ddat, unit = "day")
```

```{r}
 ddat <- mdy_hms("02/14/1970 24:00:00")
 round_date(ddat, unit = "month")
```

# **Three Different Types of Time Spans**

## **Durations - Keeping Track of Seconds**

> -   A `duration` counts the exact number of seconds in a time span, between two dates.
>
> Durations allow exact comparisons with other durations.
>
> -   However, they don't always match the number of hours, months or days you might calculate.
>
> -   We can find out [Lady Gaga's](https://en.wikipedia.org/wiki/Lady_Gaga) "approximate" age using durations

```{r}
d1 <- ymd(19860328)
d2 <- today()
agesec <- as.duration(d2 - d1)
agesec
```

> Create durations from years with `dyears()`, from days with `ddays()`, etc...

```{r}
dyears(1)
```

```{r}
ddays(1)
```

```{r}
dhours(1)
```

```{r}
dminutes(1)
```

```{r}
dseconds(1)
```

> You can add durations to date-times, but you always add *seconds*, so if there is a daylight savings change, you get weird results (add a day but the time is not the same as the time the previous day).

```{r}
one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm
```

```{r}
one_pm + ddays(1)
```

## **Periods**

> Periods track the change in the human "clock time" between two date-times by ignoring actual physical time spans.
>
> -   Adding a `period` takes into account daylight savings.
>
> be careful about the functions: remember which ones are for extracting, which ones are for durations, and which ones are for periods

```{r}
one_pm
```

```{r}
one_pm + days(1)
```

## **Intervals**

> -   Intervals record time spans as a *sequence of seconds* beginning at a specified date.
>
> -   Intervals are like durations, but with an associated start time.
>
> -   Intervals and durations are good for physical processes that don't care about human clock time
>
> -   You can read more about intervals with

```{r}
help("Interval-class")
```

> -   Working with intervals requires some care:
>
>     -   **Divide an interval by a duration to determine its physical length**,
>
>     -   **Divide an interval by a period to determine its *implied length* in clock time**.

# **Time Zones**

> -   Time zones are specified using the `tz` or `tzone` arguments (for example, in the call to `ymd_hms()` above).
>
> -   Time zones are specified by "content/city." For example, `"America/New_York"` and `"Europe_Paris"`
>
> -   You can see a complete list of time zones with `OlsonNames()`.
>
> -   The default time zone is `UTC` (which has no daylight savings). Coordinated Universal Time \*UTC) is the successor time standard to Greenwich Mean Time (GMT) which is a specific time zone.
>
> -   You usually don't have to worry about timezones unless they loaded incorrectly. For example, R might think it's `UTC` even though it should be `America/New_York` and then forget daylight savings.
>
> -   If a date-time is labeled with the incorrect time zone, use `force_tz()`

```{r}
d1 <- ymd_hms("20140101 10:01:11")
d1
```
