---
title: "class_10/31_dbplyr_stringr"
author: "lisa liubovich"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

> What is a database?
>
> When you have a huge collection of data, you want to store it efficiently. You would have a relational table (which is essentially what a data base is). You can have anything inside the cells. A database is the most efficient way to store huge data.
>
> You would send queries (anything written in the code) through SQL. Once you have a SQL query, the database management system (dbMS) does the job.
>
> If you want to work with a database, you need to use SQL. But since we don't know it, the {dbplyr} package is how we are going to work with a database. dbplyr connects to rsqlite, which is the backend package that SQL uses and which converts our R code to SQL queries.
>
> See other backend packages in the lecture notes.

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("dbplyr")
library(dbplyr)
install.packages("RSQLite")
library(RSQLite)
```

```{r}
con <- dbConnect(drv = SQLite(), dbname = "./data/soccer.sqlite")
dbListTables(con)
```

```{r}
# these are lists, not dataframes. but we can treat them as dfs
Team_db <- tbl(con, "Team")
Team_at_db <- tbl(con, "Team_Attributes")
Country_db <- tbl(con, "Country")
League_db <- tbl(con, "League")
Match_db <- tbl(con, "Match")
```

> we can now use dplyr to interact with these data frames almost like they were actually in memory.
>
> it may be easier to work in df format, so use the collect function.

> Once you select the variables you want and the observations you want, you should use collect() to get the data frame into memory so that you can have all of the functionality of R (e.g., gather() and spread() will only work on in-memory data frames). see example below:

```{r}
Match_db %>%
  select(id:away_team_goal) %>%
  collect() ->
  Match
Team_db %>%
  collect() ->
  Team
Country_db %>%
  collect() ->
  Country
```

> The following will return a data frame telling you where each team is from.

```{r}
Match %>%
  select(country_id, home_team_api_id, away_team_api_id) %>%
  gather(-country_id, key = "home_away", value = "team_api_id") %>%
  select(-home_away) %>%
  distinct() %>%
  left_join(Team, by = "team_api_id") %>%
  left_join(Country, by = c("country_id" = "id")) %>%
  select(team_long_name, team_short_name, name) %>%
  rename(country_name = name)
```

**Exercise 4.3.1 (possible quiz question!!!!)**

## **Stringr**

> Strings: In R, strings (also called "characters") are created and displayed within quotes
>
> We want to work with a sequence of strings...

```{r}
x <- "he said: \"Lisa is a good student\""
writeLines(x)
```

> the backslashes are necessary in order to get the specific output that we want. This is called an "escape" character. writeLines() is better than print in this scenario because it prints the string itself as opposed to the printed representation of the string that we would get with print().
>
> other escape characters:

```{r \n is new line}
x <- "Not all those\nwho wonder are lost."
writeLines(x)
```

```{r \t is tab}
x <- "Not all those\twho wonder are lost."
writeLines(x)
```

> You can add any Unicode character with a \\u followed by the hexadecimal unicode representation of that character --\> good for Greek letters and special characters

```{r}
mu <- "\u00b5"
writeLines(mu)
```

> The stringr package contains a lot of convenience functions for manipulating strings (and they are a lot more user friendly than base R's string manipulation functions like grep() and gsub()).
>
> stingr is part of the tidyverse so you do not have to load it separately.
>
> All of stringr's functions begin with "str\_", so you can press tab after typing "str\_" and a list of possible string manipulation functions will pop up (in RStudio)
>
> For example, to get the number of characters in a string, use str_length(). See example below

```{r}
str_length("Upon the hearth the fire is red,")
```

> What about spaces and punctuation? They count. What about escaped charaters? The \\ does but the character does not.

```{r}
str_length(x <- "Upon the hearth \nthe fire ir red,")
```

> Combine strings with str_c().

```{r}
x <- "Faithless is he that says"
y <- "farewell when the road darkens."
str_c(x, y)
```

> see that we have an issue: the strings are not separated. The default is to separate strings by nothing. Use sep to change the separator.

```{r}
str_c(x, y, sep = " ")
```

> Just like c(), str_c() can take multiple arguments.

```{r}
str_c("Short", "cuts", "make", "long", "delays.", sep = " ")
```

> If you provide str_c() a vector of arguments, it will vectorize the combining unless you provide a collapse argument.

```{r}
x <- c("Short", "cuts", "make", "long", "delays.") # vector of length 5
str_c(x, "LOTR", sep = " ") # combining a vector of length 5 with a vector of length 1
```

> *remember with vectors*: c(1,2) + c(3,4,5,6) is c(4,6,6,8) because 1+3 is 4, 2+4 is 6, 1+5 is 6, and 2+6 is 8 (the values of the first vector are added to the first two values of the second vector AND THEN the last two values of the second vector --\> if we're combining vectors with two different sizes, the output takes the length of the longer vector. the length matters, not the values.

> besides separator (sep), str_c() also has an argument called collapse. If collapse = NULL (the default) a character vector with length equal to the longest input. If collapse is a string, a character vector of length 1.

```{r}
x <- c("1","2")
str_c(x, "a", sep = ";", collapse = "???")
```

> this collapses the vector based on the argument we specified. So instead of it being "1:a", "2;a", the output is "1;a???2;a".

> Combining with NA results in NA:

```{r}
str_c("Faithless is he that says", NA)
```

> combining something known ("faithless is he that says" with something unknown (NA) logically outputs the unknown.

> str_sub() extracts a substring between the location of two characters.

```{r}
x <- "The Road goes ever on and on"
str_sub(x, start = 2, end = 6)
```

> Replace substrings with assignment

```{r}
str_sub(x, start = 2, end = 6) <- " Tolkein "
x
```
