---
title: "hw_02"
author: "lisa liubovich"
date: "2023-09-20"
output: pdf_document
---

HW 02

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

**Question 1**

a\.

1.  the function checks if the input is divisible by 5
2.  if the input is divisible by 5, the function returns "fizz"
3.  if the input is not divisible by 5, check if the input is divisible by 3
4.  if the input is divisible by 3, the function returns "buzz"
5.  the function checks if the input is divisible by both 3 and 5
6.  if the input is divisible by both 3 and 5, the function returns "buzzfizz"
7.  if the input is neither divisible by 3 or 5, the function returns the input number

b\.

```{r step 1; check if divisible by 5}
x<- 3
x %% 5 == 0

x<- 5
x %% 5 == 0

x<- 15
x %% 5 == 0

x<- 16
x %% 5 == 0

```

```{r step 2: if divisible by 5, return "fizz"}
x<- 3
if (x %% 5 == 0) print("fizz")

x<- 5
if (x %% 5 == 0) print("fizz")

x<- 15
if (x %% 5 == 0) print("fizz")

x<- 16
if (x %% 5 == 0) print("fizz")

```

```{r step 3; if not divisible by 5, check if divisible by 3}
x<- 3
x %% 3 == 0

x<- 5
x %% 3 == 0

x<- 15
x %% 3 == 0

x<- 16
x %% 3 == 0


```

```{r step 4; if input is divisible by 3, return "buzz"}
x<- 3
if (x %% 3 == 0) print("buzz")

x<- 5
if (x %% 3 == 0) print("buzz")

x<- 15
if (x %% 3 == 0) print("buzz")

x<- 16
if (x %% 3 == 0) print("buzz")
```

```{r step 5; check to see if divisible by both}
x<- 3
x %% 3 == 0 & x %% 5 == 0

x<- 5
x %% 3 == 0 & x %% 5 == 0

x<- 15
x %% 3 == 0 & x %% 5 == 0

x<- 16
x %% 3 == 0 & x %% 5 == 0
```

```{r step 6; if divisible by 3 and 5, return "buzzfizz"}
x<- 3
if (x %% 3 == 0 & x %% 5 == 0) 
  print("buzzfizz")

x<- 5
if (x %% 3 == 0 & x %% 5 == 0) 
  print("buzzfizz")

x<- 15
if (x %% 3 == 0 & x %% 5 == 0) 
  print("buzzfizz")

x<- 16
if (x %% 3 == 0 & x %% 5 == 0) 
  print("buzzfizz")
```

```{r step 7; if not divisible by 3 or 5, return input}
x<- 3
if (!(x %% 3 == 0 || x %% 5 == 0))
  print(x)

x<- 5
if (!(x %% 3 == 0 || x %% 5 == 0))
  print(x)

x<- 15
if (!(x %% 3 == 0 || x %% 5 == 0))
  print(x)

x<- 16
if (!(x %% 3 == 0 || x %% 5 == 0))
  print(x)
```

c\.

```{r final function with default argument}
buzzfizz <- function(x) {
  if (is.null(x)) {
    print("no input provided. using default value of 0.")
    x <- 0
  }
  
  if (x %% 3 == 0 && x %% 5 == 0) {
    print("buzzfizz")
  } else if (x %% 3 == 0) {
    print("buzz")
  } else if (x %% 5 == 0) {
    print("fizz")
  } else {
    print(x)
  }
}
```

d\.

```{r function tested on values}
buzzfizz(3)
buzzfizz(5)
buzzfizz(15)
buzzfizz(2)
```

e\.

```{r, error=TRUE updated function with error checking}
buzzfizz <- function(x) {
  # check to see if if x is numeric and a single value
  if (!is.numeric(x) || length(x) != 1) {
    stop("Input must be a numeric single value.")
  }
  
  if (is.null(x)) {
    warning("No input provided. Using default value of 0.")
    x <- 0
  }
  
  if (x %% 3 == 0 && x %% 5 == 0) {
    print("buzzfizz")
  } else if (x %% 3 == 0) {
    print("buzz")
  } else if (x %% 5 == 0) {
    print("fizz")
  } else {
    print(x)
  }
}

Testing with invalid inputs:
Error: Input must be a numeric single value.
Error: Input must be a numeric single value.

Testing with valid inputs:
Result 3: buzzfizz
Result 4: 7

```

f\.

```{r}
#' BuzzFizz Function
#'
#' This function checks the divisibility of a numeric input value \code{x} and returns one of the following:
#' - "buzzfizz" if \code{x} is divisible by both 3 and 5.
#' - "buzz" if \code{x} is divisible by 3.
#' - "fizz" if \code{x} is divisible by 5.
#' - The input \code{x} itself if it's not divisible by 3, 5, or both.
#'
#' @param x Numeric value to be checked for divisibility.
#'
#' @return A character string indicating divisibility or the input value itself.
#'
#' @details This function also includes error checking to ensure that the input \code{x} is a numeric single value.
#' If \code{x} is not numeric or not a single value, an error will be raised.
#'
#' @examples
#' # Test with valid inputs
#' buzzfizz(15) # Returns "buzzfizz"
#' buzzfizz(7)  # Returns 7
#'
#' # Test with invalid inputs
#' tryCatch(
#'   {
#'     buzzfizz("cat")   # Error: Input must be a numeric single value.
#'     buzzfizz(c(1, 5)) # Error: Input must be a numeric single value.
#'   },
#'   error = function(e) {
#'     cat("Error:", e$message, "\n")
#'   }
#' )
#'
#' @export
buzzfizz <- function(x) {
  # Check if x is numeric and a single value
  if (!is.numeric(x) || length(x) != 1) {
    stop("Input must be a numeric single value.")
  }
  
  if (is.null(x)) {
    warning("No input provided. Using default value of 0.")
    x <- 0
  }
  
  if (x %% 3 == 0 && x %% 5 == 0) {
    print("buzzfizz")
  } else if (x %% 3 == 0) {
    print("buzz")
  } else if (x %% 5 == 0) {
    print("fizz")
  } else {
    print(x)
  }
}

```

**Question 2**

```{r}
set.seed(1)

result <- 1:10 %>%
  sample(1000, replace = TRUE) %>%
  mean() %>%
  exp()

cat("Result:", result, "\n")


```

**Question 3**

```{r}
set.seed(1)

sample_values <- rnorm(100, mean = 10, sd = sqrt(3))

proportion_greater_than_12 <- sum(sample_values > 12) / length(sample_values)

cat("Proportion greater than 12:", proportion_greater_than_12, "\n")

```

**Question 4**

```{r}
x <- c(TRUE, FALSE, TRUE, TRUE)
y <- c(FALSE, FALSE, TRUE, FALSE)
z <- NA

result1 <- x & y
result2 <- x & z
result3 <- !(x | y)
result4 <- x | y
result5 <- y | z
result6 <- x[y]
result7 <- y[x]
result8 <- x[x | y]

cat("x & y:", result1, "\n")
cat("x & z:", result2, "\n")
cat("!(x | y):", result3, "\n")
cat("x | y:", result4, "\n")
cat("y | z:", result5, "\n")
cat("x[y]:", result6, "\n")
cat("y[x]:", result7, "\n")
cat("x[x|y]:", result8, "\n")

```
