---
title: "hw_03"
author: "lisa liubovich"
date: "2023-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1**

```{r}
options(repos = "https://cran.r-project.org")
install.packages("tidyverse")
library(tidyverse)
install.packages("ggplot2")
library(ggplot2)
data(msleep)
install.packages("ggthemes")
library(ggthemes)
```

**Question 2**

```{r}
nrow(msleep)
ncol(msleep)

```

> There are 83 mammals and 11 variables in the msleep data frame.

**Question 3**

| Does change in body weight explain change in total sleep time?
| Explanatory variable: body weight, (numerical, continuous)

```{r}
ggplot(msleep, mapping = aes(x = bodywt, y = sleep_total)) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess) +
  xlab("Body Weight") +
  ylab("Total Sleep Time") +
  ggtitle("Body Weight vs Total Sleep Time")
```

> While perhaps there is some kind of a trend to the data, the shape of the plot indicates that there is no correlation between total sleep time and body weight.

**Question 4**

> a\.

```{r plot 1}
ggplot(msleep, mapping = aes(x = log(bodywt), y = sleep_total)) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess) +
  xlab("Log of Body Weight") +
  ylab("Total Sleep Time") +
  ggtitle("Log of Body Weight vs Total Sleep Time")
```

```{r plot 2}
ggplot(msleep, mapping = aes(x = bodywt, y = log(sleep_total))) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess) +
  xlab("Body Weight") +
  ylab("Log of Total Sleep Time") +
  ggtitle("Body Weight vs Log of Total Sleep Time")
```

```{r plot 3}
ggplot(msleep, mapping = aes(x = log(bodywt), y = log(sleep_total))) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess) +
  xlab("Log of Body Weight") +
  ylab("Log of Total Sleep Time") +
  ggtitle("Body Weight vs Log of Total Sleep Time")
```

> b\. I would say that plot one is the most linear because it appears to be the closest to a straight line while plots 2 and 3 have almost parabola-like shapes.

> c\.

```{r}
ggplot(msleep, mapping = aes(x = bodywt, y = sleep_total)) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess) +
  scale_x_log10() +
  xlab("Body Weight") +
  ylab("Total Sleep Time") +
  ggtitle("Body Weight vs. Total Sleep Time")
```

> d.  This plot is exactly the same in shape as my best plot (plot 1). The main difference here is the scale; my best plot's x-axis represents the logarithm of body weight while the new plot's x-axis represents powers of 10 of body weight.

**Question 5**

```{r}
ggplot(msleep, mapping = aes(x = log(bodywt), y = sleep_total, color = vore)) +
  geom_point() +
  geom_smooth(se = TRUE, method = loess, color = "black") +
  xlab("Log of Body Weight (ln(kg)") +
  ylab("Total Sleep Time (hours)") +
  ggtitle("Log of Body Weight vs. Total Sleep Time, colored by diet") +
  theme_bw()
```

**Question 6**

```{r}
ggplot(msleep, mapping = aes(x = log(bodywt), y = sleep_total, color = vore)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm) +
  xlab("Log of Body Weight (ln(kg)") +
  ylab("Total Sleep Time (hours)") +
  ggtitle("Log of Body Weight vs. Total Sleep Time, colored by diet") +
  theme_bw() +
  facet_wrap(~ vore)
```

> The effect of body weight on sleep total appears to be larger for herbivores and carnivores compared to insectivores, omnivores, and mammals with unspecified diets.

**Question 7**

```{r}
ggplot(msleep, mapping = aes(x = log(bodywt), y = sleep_total, color = vore)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm, color = "black", linetype = "dashed", size = 0.5) +
  xlab("Log of Body Weight (ln(kg)") +
  ylab("Total Sleep Time (hours)") +
  ggtitle("Log of Body Weight vs. Total Sleep Time, colored by diet") +
  theme_bw()
```

> This line shows a stronger negative relationship between body weight and total sleep time for omnivorous, insectivorous, and unspecificied diets, but a weaker negative relationship between body weight and total sleep time for carnivorous and herbivorous diets.

**Question 8**

```{r}
ggplot(msleep, mapping = aes(x = log(bodywt), y = sleep_total, color = vore)) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm, color = "black", linetype = "dashed", size = 0.5) +
  xlab("Log of Body Weight (ln(kg)") +
  ylab("Total Sleep Time (hours)") +
  ggtitle("Log of Body Weight vs. Total Sleep Time, colored by diet") +
  theme_bw() +
  guides(color = guide_legend(title = "Diets"))
```

**Question 9**

```{r}
ggplot(msleep, mapping = aes(x = vore, y = sleep_total, fill = vore, color = vore)) +
  geom_boxplot(outlier.shape = 17, color = "gray20") +
  scale_y_log10() +
  xlab("Diet") +
  ylab("Sleep Total") +
  scale_fill_colorblind() +
  theme_bw() +
  theme(
    legend.position = "none"
    )
```

**Question 10**

```{r}
ggplot(msleep, mapping = aes (x = bodywt, y = sleep_total, fill = "black")) +
  geom_point() +
  geom_smooth(se = FALSE, method = lm, color = "darkgreen") +
  xlab("Body Weight") +
  ylab("Sleep Total") +
  scale_x_log10() +
  scale_y_log10() +
  theme_bw() +
  ggtitle("Log(Sleep Total) vs Log(Body Weight)") +
  facet_wrap(~ vore) +
  theme(
    legend.position = "none"
  )
```
