---
title: "test_1"
author: "lisa liubovich"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.r-project.org"))
```

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("ggplot2")
library(ggplot2)
install.packages("dplyr")
library(dplyr)
```

**Question 1**

```{r}
# piping method
midwest %>% 
  filter(state %in% c("OH"))
```

```{r}
# assignment method
ohio_data <- filter(midwest, state == "OH")
head(ohio_data)
```

**Question 2**

```{r}
midwest %>% 
  filter((state == "OH"),
         popwhite > 50000,
         popwhite < 90000)
```

**Question 3**

```{r}
midwest %>% 
  select(state:county, poptotal, popamerindian, percamerindan) %>% 
  arrange(desc(popamerindian)) %>% 
  slice_head(n = 20)
```

**Question 4**

```{r}
midwest %>% 
  filter((state == "MI"),
         (poppovertyknown > 10000 & percprof > 10)) %>% 
  select(state:county, poppovertyknown, percprof) 
```

**Question 5**

```{r}
midwest %>% 
  mutate(mean_poverty_pop = mean(poppovertyknown)) %>% 
  group_by(county) %>% 
  select(county:state, mean_poverty_pop)
```

**Question 6**

```{r}
ggplot(data = midwest, mapping = aes(x = poptotal, y = poppovertyknown)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Total Population") +
  ylab("Known Poverty Population") +
  ggtitle("Relationship Between Total Population and Known Poverty Population")
```

**Question 7**

```{r}
ggplot(data = midwest, mapping = aes(x = state, y = percadultpoverty)) +
  geom_boxplot() +
  xlab("state") +
  ylab("percadultpoverty")
```

**Question 8**

```{r}
ggplot(data = midwest, mapping = aes(x = percollege, y = percprof)) +
  geom_point(color = "red") +
  geom_smooth(se = FALSE, method = lm, color = "black") +
  xlab("percollege") +
  ylab("percprof") +
  ggtitle("College/Professional Work Scatterplots") +
  facet_wrap(.~state)
```

**Question 9**

> Illinois has the highest count with slightly over 100.

```{r}
ggplot(data = midwest, mapping = aes(x = state, fill = state)) +
  geom_bar()
```

**Question 10**

> part a: the str() command shows the structure of the dataset midwest. The first line of the output (tibble [437 x 28] .....) confirms that midwest is a tibble.

```{r}
str(midwest)
```

> part b: the first method subsets poptotal and shows the output as a vector. The second method subsets poptotal and shows the output as a tibble.

```{r}
# first method
midwest$poptotal
```

```{r}
subset <- midwest["poptotal"]
print(subset)
```

> part c:

```{r}
midwest %>% 
  group_by(state) %>% 
  summarize(mean_area = mean(area)) %>% 
  select(state, mean_area)
```

**Question 11**

```{r}
midwest %>% 
  mutate(Ratio = popadults/poptotal) %>% 
  mutate(Percent = Ratio*100) %>% 
  select(state, county, poptotal, popadults, Ratio, Percent) %>% 
  filter(state == "WI") %>% 
  slice_head(n = 25)
```

**Question 12**

```{r}
ggplot(data = midwest, mapping = aes(x = area, y = state, fill = state)) +
  geom_violin() +
  ggtitle("Violin Plots (area vs state)")
```
